---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# TRNValStandVis

<!-- badges: start -->
<!-- badges: end -->

## Description

`TRNValStandVis` is an R packacge that validates and 
It combines transcription factor - target gene interaction data from four data sets, 
the Molecular Signatures Database (MSigDB) by GSEA, hTFtarget by Zhang et al.,
the ChIP-X Enrichment Analysis Version 3 (ChEA3) by Ma'ayan lab, and TFLink by Liska et al. to 
create a single combined dataset, where all TF's and TG's are standardized to be HGNC compliant. 
The combined data was used to create a qualitative score to evaluate the quality of TF-TG interactions
as well a score to assess the overall quality of a specific TF. Further abilities include
visualizing TF-TG regulons and networks and comparing different TF's qualitatively. 
This package will allow for the validation of TF-TG's across different sources, 
quality metrics in regards to level of validation for each TF-TG, as well as 
standardization in regards to nomenclature. The `TRNValStandVis` package was developed using
`R version 4.1.1 (2021-08-10)`, `Platform: x86_64-apple-darwin17.0 (64-bit)` and `Running under: macOS Big Sur 11.6`


## Installation

In order to install the latest version of this package:

```{r, eval=FALSE}
install.packages("devtools")
library("devtools")
devtools::install_github("Mogtaba-Alim/TRNValStandVis", build_vignettes = TRUE)
library("TRNValStandVis")
```

## Overview

```{r, eval=FALSE}
ls("package:TRNValStandVis")
data(package = "TRNValStandVis")
browseVignettes("TRNValStandVis")
```

`TRNValStandVis` consists of 8 functions. 

1. ***get_TGs***
- Take in a TF and return all its target genes. Can be adjusted based on number of datasets TF-TG appears in.

2. ***get_TFs***
- Take in a gene and return all the transcription factors that regulate it. Can be adjusted based on number of datasets TF-TG appears in. 

3. ***plot_TF_regulon***
- Take in a gene and plot that gene's regulon including regulators and targets. 

4. ***plot_TFs_network***
- Take in up to 3 gene names and plot the network involving all the genes regulons combined. 

5. ***get_TF_TG_val***
- Take in a TF and return the validity score for all the TG's of that TF

6. ***plot_TF_TG_val***
- Take in a TF and plot the validity score for all the TG's of that TF

7. ***get_TF_total_val***
- Takes in a TF and provides the average validty score for that TF

8. ***plot_TF_total_val***
- Take in a collection of TF names and plots their average validity score

The packcage will also have 2 datasets. The first  will be the combined dataset of TF-TG's
and will contain columns for the number of times an interaction has appeared in the
different datasets, the validation score for each TF-TG and a normalized validation score. 

The second dataset, will be of all the TF's and their overall average validation score. 

## Contributions

The author of this package is Mogtaba Alim. All functions and processes including
the processing and the creation of valdity scores were done by the author without any 
external help. 

I used the following three external packages, igraph, ggplot2 and igraph
throughout the package. 

## References

Dplyr. (n.d.). Retrieved November 15, 2023, from Tidyverse.org website: https://dplyr.tidyverse.org/

Fujiyoshi, K., Bruford, E. A., Mroz, P., Sims, C. L., O’Leary, T. J., Lo, A. W. I., … Ogino, S. (2021). Standardizing gene product nomenclature—a call to action. Proceedings of the National Academy of Sciences of the United States of America, 118(3). doi:10.1073/pnas.2025207118

Ggplot2. (n.d.). Retrieved November 15, 2023, from Tidyverse.org website: https://ggplot2.tidyverse.org

igraph – Network analysis software. (n.d.). Retrieved November 15, 2023, from Igraph.org website: https://igraph.org/

Lachmann, A., Xu, H., Krishnan, J., Berger, S. I., Mazloom, A. R., & Ma’ayan, A. (2010). ChEA: transcription factor regulation inferred from integrating genome-wide ChIP-X experiments. Bioinformatics (Oxford, England), 26(19), 2438–2444. doi:10.1093/bioinformatics/btq466

Liberzon, A., Birger, C., Thorvaldsdóttir, H., Ghandi, M., Mesirov, J. P., & Tamayo, P. (2015). The molecular signatures database hallmark gene set collection. Cell Systems, 1(6), 417–425. doi:10.1016/j.cels.2015.12.004

Liska, O., Bohár, B., Hidas, A., Korcsmáros, T., Papp, B., Fazekas, D., & Ari, E. (2022). TFLink: an integrated gateway to access transcription factor–target gene interactions for multiple species. Database: The Journal of Biological Databases and Curation, 2022, baac083. doi:10.1093/database/baac083

The R project for statistical computing. (n.d.). Retrieved November 15, 2023, from R-project.org website: http://www.R-project.org

Zhang, Q., Liu, W., Zhang, H.-M., Xie, G.-Y., Miao, Y.-R., Xia, M., & Guo, A.-Y. (2020). HTFtarget: A comprehensive database for regulations of human transcription factors and their targets. Genomics, Proteomics & Bioinformatics, 18(2), 120–128. doi:10.1016/j.gpb.2019.09.006

## Acknowledgements

This package was developed as part of an assessment for 2023 BCB410H: Applied Bioinformat-
ics course at the University of Toronto, Toronto, CANADA. `TRNValStandVis` welcomes issues,
enhancement requests, and other contributions. To submit an issue, use the GitHub issues.

